SSwAB-IT Phase 1 Analysis

### 2024-03-06 Jordana Hoppe/SSwAB-IT/phase_1_analysis.qmd
Update phase_1_analysis.qmd

---
title: "SSwAB-IT Phase 1 Analysis"
author: "Tim Vigers"
date: "today"
date-format: long
format:
  html:
    toc: true
    toc-depth: 5
    toc-float: true
    code-fold: true
    self-contained: true
    fig-cap-location: top
    page-layout: full
    theme:
      light: flatly
      dark: darkly
bibliography: /Users/timvigers/Documents/Miscellaneous/zotero.bib
csl: /Users/timvigers/Documents/Miscellaneous/american-medical-association.csl
editor: source
---

@@

```{r data cleaning}
#| include: false
# Import from REDCap
unlockREDCap(c(rcon = "SSwAB-IT (New)"),
  keyring = "API_KEYs",
  envir = 1,
  url = "https://redcap.ucdenver.edu/api/"
)
exportBulkRecords(list(db = rcon), envir = 1)
# Combine collection results
df <- full_join(db_cos_collection_results, db_sos_collection_results)
# Format
df <- df %>%
  select(
    record_id, redcap_data_access_group,
    staphylococcus_aureus_cos:other_result_cos,
    staphylococcus_aureus_sos:other_result_sos
  )
```

@@

```{r}
agreement <- function(bug, data = df) {
  cos <- paste0(bug, "_cos")
  sos <- paste0(bug, "_sos")
  d <- data[, c(cos, sos)]
  a <- agree(d)
  k <- kappa2(d)
  ct <- table(d[, cos])
  st <- table(d[, sos])
  return(list(
    "COS Absent" = ct[1], "COS Present" = ct[2],
    "SOS Absent" = st[1], "SOS Present" = st[2],
    "Percent Agreement" = a$value, "Cohen's Kappa" = k$value,
    "p value" = k$p.value
  ))
}
```

@@

@@

```{r}
bugs <- c(
  "staphylococcus_aureus", "haemophilus_influenza", "pseudomonas_aeruginosa",
  "pseudomonas_non_aero", "s_maltophilia", "burkholderia", "aspergillus", "ntm",
  "other"
)
bug_names <- c(
  "S. aureus", "H. influenza", "P. aeruginosa", "Pseudomonas non-aeruginosa",
  "S. maltophilia", "Burkholderia", "Aspergillus", "NTM", "Other"
)
bug_list = as.list(bug_names)
names(bug_list) = bugs
l <- lapply(bugs, function(b) {
  agreement(b)
})
l <- data.frame(do.call(rbind, l), check.names = F)
l <- data.frame(lapply(l, as.numeric), check.names = F)
rownames(l) <- bug_names
kable(l, digits = 2)
```

```{r results='asis'}
invisible(lapply(names(bug_list),function(b){
  cat("\n")
  cat("##",as.character(bug_list[b]))
  cat("\n")
  cos <- paste0(b, "_cos")
  sos <- paste0(b, "_sos")
  d <- df[, c(cos, sos)]
  t = data.frame(table(d))
  colnames(t) = c("COS","SOS","n")
  cat("\n")
  print(kable(t))
  cat("\n")
}))
```

>>>>

---
title: "SSwAB-IT Phase 1 Analysis"
author: "Tim Vigers"
date: "today"
date-format: long
format:
  html:
    toc: true
    toc-depth: 5
    toc-float: true
    code-fold: true
    self-contained: true
    fig-cap-location: top
    page-layout: full
    theme:
      light: flatly
      dark: darkly
bibliography: /Volumes/Documents/Miscellaneous/zotero.bib
csl: /Volumes/Documents/Miscellaneous/american-medical-association.csl
editor: source
---

@@

```{r data cleaning}
#| include: false
# Import from REDCap
unlockREDCap(c(rcon = "SSwAB-IT (New)"),
  keyring = "API_KEYs",
  envir = 1,
  url = "https://redcap.ucdenver.edu/api/"
)
exportBulkRecords(list(db = rcon), envir = 1)
# Exclude
exclude <- c("01-030_W", "01-038_W", "02-0038", "02-0039", "04-001")
# Combine collection results
df <- full_join(db_cos_collection_results, db_sos_collection_results)
# Format
df <- df %>%
  select(
    record_id, redcap_data_access_group,
    staphylococcus_aureus_cos:other_result_cos,
    staphylococcus_aureus_sos:other_result_sos
  ) %>%
  filter(!record_id %in% exclude)
# Remove empty demographics rows
db_demographics <- db_demographics %>%
  filter(
    redcap_event_name == "Clinic Visit (Arm 1: Eligible)",
    !record_id %in% exclude
  )
# Excluded participants
db_medications <- db_medications %>% filter(!record_id %in% exclude)
# Combine race and ethnicity columns
races <- c(
  "American Indian or Alaskan Native", "Asian", "Hawaiian or Pacific Islander",
  "Black or African American", "White", "Unknown", "Other"
)
ethnicities <- c("Hispanic", "Non-Hispanic", "Unknown/Not Reported")
db_demographics$Race <- apply(db_demographics, 1, function(r) {
  w <- which(r[grep("race", names(r))] == "Checked")
  if (length(w) == 1) {
    return(races[w])
  } else if (length(w) > 1) {
    return("Multiple Races")
  } else if (length(w) == 0) {
    return(NA)
  }
})
db_demographics$Ethnicity <- apply(db_demographics, 1, function(r) {
  w <- which(r[grep("ethnicity", names(r))] == "Checked")
  if (length(w) == 1) {
    return(ethnicities[w])
  } else if (length(w) > 1) {
    return(paste(ethnicities[w], sep = "/"))
  } else if (length(w) == 0) {
    return(NA)
  }
})
# # Drop levels
# level_drop <- c("sex", "Race", "Ethnicity")
# db_demographics[, level_drop] <- lapply(
#   db_demographics[, level_drop],
#   function(c) {
#     droplevels(factor(c))
#   }
# )
# Add groups to other dataframes
db_medications <- left_join(db_medications,
  db_demographics %>% select(record_id, group),
  by = join_by(record_id)
)
```

# Data cleaning

Participants 01-030_W, 01-038_W, 02-0038, 02-0039, and 04-001 were excluded from all analyses. Participants missing either the COS or SOS sample were excluded from agreement analyses, but retained in feasibility and other analyses.

# Demographics

```{r}
db_demographics %>%
  select(group, sex, Race, Ethnicity) %>%
  tbl_summary(by = group, label = list("sex" = "Sex")) %>%
  add_overall() %>%
  add_p(test = list(c("Race", "Ethnicity") ~ "fisher.test")) %>%
  separate_p_footnotes() %>%
  as_gt()
```

# Medications

```{r}
db_medications %>%
  select(group, azithromycin:please_specify_mucolytic) %>%
  select(
    -inhaled_colisitin_status, -inhaled_vancomycin_status,
    -other_inhaled_status, -please_specify_mucolytic
  ) %>%
  tbl_summary(by = group) %>%
  add_overall() %>%
  add_p() %>%
  separate_p_footnotes() %>%
  as_gt()
```

@@

```{r}
agreement <- function(bug, data = df) {
  cos <- paste0(bug, "_cos")
  sos <- paste0(bug, "_sos")
  d <- data[, c(cos, sos)]
  d <- d[complete.cases(d), ]
  a <- agree(d)
  k <- kappa2(d)
  ct <- table(d[, cos])
  st <- table(d[, sos])
  return(list(
    "COS Absent" = ct[1], "COS Present" = ct[2],
    "SOS Absent" = st[1], "SOS Present" = st[2],
    "Percent Agreement" = a$value, "Cohen's Kappa" = k$value,
    "p value" = k$p.value
  ))
}
```

@@

To calculate overall agreement, we determined whether a given COS or SOS sample was positive for *any* organism, and then calculated agreement metrics.

```{r}
bugs <- c(
  "staphylococcus_aureus", "haemophilus_influenza", "pseudomonas_aeruginosa",
  "pseudomonas_non_aero", "s_maltophilia", "burkholderia", "aspergillus", "ntm",
  "other"
)
bug_names <- c(
  "S. aureus", "H. influenza", "P. aeruginosa", "Pseudomonas non-aeruginosa",
  "S. maltophilia", "Burkholderia", "Aspergillus", "NTM", "Other"
)
bug_list <- as.list(bug_names)
names(bug_list) <- bugs
# Create overall columns (any bug detected in sample)
df$overall_cos <- apply(df, 1, function(r) {
  any(r[paste0(bugs, "_cos")] == "Yes")
})
df$overall_cos <- factor(df$overall_cos, levels = c(F, T), labels = c("Negative", "Positive"))

df$overall_sos <- apply(df, 1, function(r) {
  any(r[paste0(bugs, "_sos")] == "Yes")
})
df$overall_sos <- factor(df$overall_sos, levels = c(F, T), labels = c("Negative", "Positive"))
# Agreement metrics
d <- df[, c("overall_cos", "overall_sos")]
d <- d[complete.cases(d), ]
a <- agree(d)
k <- kappa2(d)
ct <- table(d[, "overall_cos"])
st <- table(d[, "overall_sos"])
kable(data.frame(
  "COS Negative" = ct["Negative"], "COS Positive" = ct["Positive"],
  "SOS Negative" = st["Negative"], "SOS Positive" = st["Positive"],
  "Percent Agreement" = a$value, "Cohen's Kappa" = k$value,
  "p value" = k$p.value, check.names = F
), digits = 2)
t <- data.frame(table(d))
colnames(t) <- c("COS", "SOS", "n")
kable(t)
```

@@

```{r}
l <- lapply(bugs, function(b) {
  agreement(b)
})
l <- data.frame(do.call(rbind, l), check.names = F)
l <- data.frame(lapply(l, as.numeric), check.names = F)
rownames(l) <- bug_names
kable(l, digits = 2)
```

```{r results='asis'}
invisible(lapply(names(bug_list), function(b) {
  cat("\n")
  cat("##", as.character(bug_list[b]))
  cat("\n")
  cos <- paste0(b, "_cos")
  sos <- paste0(b, "_sos")
  d <- df[, c(cos, sos)]
  d <- d[complete.cases(d), ]
  t <- data.frame(table(d))
  colnames(t) <- c("COS", "SOS", "n")
  cat("\n")
  print(kable(t))
  cat("\n")
}))
```
